---
title: "Small Worlds and Large Worlds"
author: "Stuart Lee"
date: "30/01/2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

> All statistical modelling has these sames two frames: the _small world_ of the
model itselt and the _large world_ we hope to deploy the model in.

* small world = the logic and assumptions of the model
* large world = the broad context of the deployed model, the model is never a 
full representation of the large world

Baysian inference is counting and comparing possibilities.

## Example: marbles in a bag

Here's some code to create the possible conjectures for the bag of 4 marbles
containing some combination of black and white marbles.
```{r}
white_marble <- "\U25EF"
black_marble <- "\U25C9"
observed <- c(black_marble, white_marble, black_marble)

layout_bag <- function(p, n_marbles = 4) {
  stopifnot(!(p - n_marbles > 1))
  n_black <- min(p, n_marbles + 1) - 1
  n_white <- n_marbles - n_black
  c(rep(black_marble, n_black), rep(white_marble, n_white))
}

conjecture <- tibble(possibility = 1:5,
                     bag = map(possibility, layout_bag),
                     elegant_bag = map_chr(bag, ~ str_c(.x, collapse = "")))
conjecture
```

We observe the following sequence of marbles from the bag:
`r str_c(observed, collapse = "")`.  

There are `r length(conjecture$elegant_bag[[1]]) ^ length(observed)` possible
ways of choosing these 3 marbles from the bag. 

Now for each conjecture we can count
the number of ways of producing `r str_c(observed, collapse = "")`.

```{r}
number_of_ways <- function(observed, bag) {
  if (!all(observed %in% bag)) return(0L)
  counter <- 1L
  for (i in seq_along(observed)) {
    counter <- counter * sum(bag %in% observed[i])
  }
  counter
}
conjecture <- conjecture %>% 
  mutate(n_ways = map_int(bag, ~number_of_ways(observed, .x)))
conjecture
```

We have supposed that each conjecture is equally plausible, so if we observed
another `r black_marble` we could simply include add this to `r str_c(observed, collapse = "")` and rerun the `number_of_ways` function.

```{r}
conjecture %>%
  mutate(new_count = map_int(bag, 
                             ~number_of_ways(c(observed, black_marble), .x)))
```

Alternately, we could use our previously computed counts as a prior and multiply
it by the number of ways to produce `r black_marble` for each possible bag.

```{r}
conjecture <- conjecture %>%
  mutate(n_ways_new_obs = map_int(bag, ~sum(.x %in% black_marble)),
         new_count = n_ways * n_ways_new_obs)
conjecture 
```

What if we were told by the marble factory that black marbles are rare,
then for every `r conjecture$elegant_bag[4]` there are two `r conjecture$elegant_bag[3]` and three `r conjecture$elegant_bag[2]`. Then we can 
update our counts using the same logic as before:
```{r}
conjecture <- conjecture %>%
  select(-new_count) %>%
  mutate(factory_count = c(0L, 3L:1L, 0L),
         new_count = n_ways*n_ways_new_obs*factory_count)
conjecture
```

From counts to probability, the mantra of Bayes:

posterior `r "\U221D"` likelihodd * prior

In the marbles example this means the plausibility of a bag given 
the observed sequence `r str_c(observed, collapse = "")` depends on
the ways a bag can produce `r str_c(observed, collapse = "")` multiplied
by the prior plausibility of the bag.


```{r}
conjecture_probs <- conjecture %>%
  mutate(proportion_black_marble = n_ways_new_obs / max(n_ways_new_obs),
    posterior_observed = n_ways / sum(n_ways)) %>%
  select(elegant_bag, proportion_black_marble, n_ways,  posterior_observed)
conjecture_probs
```

## Building a Model

